---
title: "MFE5209 ARCH/GARCH"
author: "Wu Fuheng"
date: "Monday, October 06, 2014"
output: html_document
---
## Problem 1
This is a AR(1)/ARCH(1) model, the conditional mean and variance of $u_t$ will be __non-constant__.

Since $\mu=0.4, \phi=0.45, \omega=1$ and $\alpha_1=0.3$, we can rewrite the equation of $u_t$ as follows:

$$\begin{align*}
u_t 
&=(1-\phi)\mu + \phi u_{t-1} + h_t\\
&=(1-0.45)0.4+0.45 u_{t-1}+h_t\\
&=0.22+0.45 u_{t-1}+h_t\\
\end{align*}$$

----

>__a)__Conditional Expectation

The conditional mean of $u_t$ is

$$\begin{align*}
E[u_t|u_{t-1},...]
&= 0.22+0.45u_{t-1}+E[h_t|h_{t-1}...]\\
\end{align*}$$

According to P479 of *SDAFE*, for a ARCH(1) process $h_t$, the conditional mean $E[h_t|h_{t-1}...]$=0. Thus we have

$$\begin{align*}
E[u_t|u_{t-1},...]
&= 0.22+0.45u_{t-1}\\
\end{align*}$$

So,

$$\begin{align*}
E[u_2|u_1=1,u_0=0.2]
&= 0.22+0.45u_1\\
&=0.67
\end{align*}$$

>__b)__Conditional Variance

The conditional variance of $u_t$ is
$$\begin{align*}
Var[u_t|u_{t-1},...]
&= E[(u_t-E[u_t|u_{t-1},...])^2|u_{t-1},...]\\
&= E[(h_t)^2|u_{t-1},...]\\
&= \omega + \alpha_1 h_{t-1}^2\\
\end{align*}$$

Thus, we have
$$\begin{align*}
Var[u_2|u_1=1,u_0=0.2]
&= \omega + \alpha_1 h_{1}^2\\
&= 1 + 0.3 h_{1}^2\\
\end{align*}$$

And we also know
$$\begin{align*}
u_1-\mu=\phi (u_0 - \mu) + h_1\\
\longrightarrow 1-0.4=0.45(0.2-0.4)+ h_1\\
\longrightarrow h_1=0.51
\end{align*}$$

Therefore, $Var[u_2|u_1=1,u_0=0.2]=1+0.3*0.51^2=1.07803$

---

## Problem 2
Assume $\beta_0=0.06, \beta_1=0.35, \delta=0.22$, we can rewrite the model

$$\begin{align*}
Y_t=0.06+0.35X_t+0.22\sigma_t+h_t
\end{align*}$$

----

>__a)__Conditional Expectation

$$\begin{align*}
E[Y_t|X_t=0.1,h_{t-1}=0.6]
&=0.06+0.35*0.1+0.22\sigma_t\\
&=0.095+0.22\sigma_t\\
\end{align*}$$

>__b)__Conditional Variance

$$\begin{align*}
Var[Y_t|X_t=0.1,h_{t-1}=0.6]
&=E[(Y_t-E[Y_t|X_t=0.1,h_{t-1}=0.6])^2|X_t=0.1,h_{t-1}=0.6]\\
&=E[h_t^2|X_t=0.1,h_{t-1}=0.6]\\
&=1+0.5h_{t-1}^2\\
&=1+0.5*0.6^2\\
&=1.18
\end{align*}$$

>__c)__Is the conditional distribution of $Y_t$ given $X_t$ and $h_{t-1}$ normal? Why or why not?

Yes, it is normal.
Reason: There is only one stochastic item $\varepsilon$ in the formula which is normal. And we know the linear combination of a normal distribution is normal.

>__d)__Is the marginal distribution of $Y_t$ normal? Why or why not?

No, it isn't.
Reason: In this case, $h_t$ is a ARCH(1) process, which is of a non-Gaussian distribution with a heavy tail. So $Y_t$ is also not Gaussian/normal.

----

## Problem 3

This problem uses monthly observations of the two-month yield, that is, $Y_T$ with T 
equal to two months, in the data set Irates in the Ecdat package. The rates are 
log-transformed to stabilize the variance. To fit a GARCH model to the changes in the 
log rates, run the following R code.

```{r  message=FALSE, warning=FALSE, results='hide'}
library(fGarch)
library(Ecdat)
data(Irates)
r = as.numeric(log(Irates[,2]))
n = length(r)
lagr = r[1:(n-1)]
diffr = r[2:n] - lagr
g=garchFit(~arma(1,0)+garch(1,1),data=diffr, cond.dist = "std")
```

>__a)__ What model is being fit to the changes in $r$? Describe the model in detail.

The model is a AR(1)/GARCH(1,1) model, assuming a t-distributed errors. We are trying to use AR(1) to model the return conditional mean and GARCH(1,1) to model its conditional variance. We can write is mathematically as follows:

$$\begin{align*}
\Delta r_t=Y_t=\mu+ar_1 Y_{t-1}+a_t\\
\\
a_t=\varepsilon_t \sqrt{\omega + \alpha_1 a_{t-1}^2+ \beta_1 \sigma_{t-1}^2},  
\varepsilon_t \thicksim t(.)
\end{align*}$$

In this case, the fitted model is:
$$\begin{align}
\Delta r_t=Y_t=0.00910873+0.09508060 Y_{t-1}+a_t\\
\\
a_t=\varepsilon_t \sqrt{0.00054864 + 0.32444019 a_{t-1}^2+ 0.74026870 \sigma_{t-1}^2}\\
\\
\varepsilon_t \thicksim t(2.83060758)
\end{align}$$

The details of the fitted model are as follows:
```{r}
summary(g)
```

>__b)__ What are the estimates of the parameters of the model?

The estimates of the model parameters are as follows

<PRE>
Coefficient(s):
         mu         ar1       omega      alpha1       beta1       shape  
 0.00910873  0.09508060  0.00054864  0.32444019  0.74026870  2.83060758  
</PRE>

>__c)__ What is the estimated ACF of $\Delta r_t$ ?

```{r  message=FALSE, warning=FALSE}
acf(diffr)
```

>__d)__ What is the estimated ACF of $a_t$ ?

$a_t$ is uncorrelated, aka $ACF(a_t)=\rho_a(h)=0$ if $h\neq 0$.
```{r  message=FALSE, warning=FALSE}
acf(residuals(g))
```

>__e)__ What is the estimated ACF of ${a_t}^2$ ?

```{r  message=FALSE, warning=FALSE}
acf(residuals(g)^2)
```



