---
title: "MFE5209 Vector AutoRegressive"
author: "Wu Fuheng"
date: "Wednesday, October 08, 2014"
output: html_document
---

## Problem 1

>__a)__

```{r message=FALSE, warning=FALSE}
library(MTS)
C=matrix(c(0.8,-0.3,0.4,0.6), nrow = 2)
S=matrix(c(2,0.5,0.5,1.0), nrow = 2)
m1 = VARMAsim(300, arlags = c(1), phi = C, sigma = S)
yt = m1$series
plot(yt)
```

>__b)__

```{r message=FALSE, warning=FALSE}
ccm(yt, lags = 5)
```

>__c)__

```{r message=FALSE, warning=FALSE}
mq(yt, lag=10)
```

We can see all p-value is 0, so we can reject $H_0$ of no cross-correlations with 5% significance level.

>__d)__

```{r message=FALSE, warning=FALSE}
m2 = VARMAsim(200,malags = c(1),theta = C,sigma = S);
zt = m2$series
ccm(zt, lags = 2)
mq(zt, lag=10)
```


## Problem 2
>The file q-fdebt.txt contains the U.S. quarterly federal debts held by (a)foreign and 
>international investors, (b) federal reserve banks, and (c) the public.The data are 
>from the Federal Reserve Bank of St. Louis, from 1970 to 2012 for 171 
>observations, and not seasonally adjusted. The debts are in billions of dollars. 
>Take the log transformation and the first difference for each time series. Let zt be 
>the differenced log series.

>__a)__ Plot the time series $z_t$.

```{r message=FALSE, warning=FALSE}
library(MTS)
da=read.table("q-fdebt.txt",header=T)
debt=log(da[,3:5])
tdx=da[,1]+da[,2]/12
#MTSplot(debt,tdx);
zt=diffM(debt);
MTSplot(zt,tdx[-1])
```

>__b)__ Obtain the first five lags of sample CCMs of $z_t$.

```{r message=FALSE, warning=FALSE}
ccm(zt, lags = 5)
```

>__c)__ Use a VAR model to fit the data $z_t$, with an appropriate order selection. Justify your order choice. Interpret the fitted model.

```{r message=FALSE, warning=FALSE}
VARorder(zt)
```


```{r message=FALSE, warning=FALSE}
v=VAR(zt,p=6)
```

>__d)__ Is the fitted model adequate? Draw the conclusion using the 5% significance level.

```{r message=FALSE, warning=FALSE}
resi=v$residuals
mq(resi,adj=18)
```


