---
title: "mfe5209hw8"
author: "Wu Fuheng"
date: "Saturday, October 11, 2014"
output: html_document
---

>There are 1352 yield curves in the data, so you will see a representative sample of them. The yield 
>curves change slowly, which is why one should look at yield curves that are spaced rather far (100 days) apart.

```{r message=FALSE, warning=FALSE}
yieldDat = read.table("9_treasury_yields.txt",header=T)
maturity = c((0:5),5.5,6.5,7.5,8.5,9.5)
pairs(yieldDat)
par(mfrow=c(4,3))
for (i in 0:11){
  plot(maturity,yieldDat[100*i+1,],type="b")
}
```

>Next compute the eigenvalues and eigenvectors of the sample covariance matrix, print the results, 
and plot the eigenvalues as a scree plot.

```{r message=FALSE, warning=FALSE}
eig = eigen(cov(yieldDat))
eig$values
eig$vectors
par(mfrow=c(1,1))
barplot(eig$values)
The following R code plots the first four eigenvectors.
par(mfrow=c(2,2))
plot(eig$vector[,1],ylim=c(-.7,.7),type="b")
abline(h=0)
plot(eig$vector[,2],ylim=c(-.7,.7),type="b")
abline(h=0)
plot(eig$vector[,3],ylim=c(-.7,.7),type="b")
abline(h=0)
plot(eig$vector[,4],ylim=c(-.7,.7),type="b")
abline(h=0)
```

>1.  It is generally recommended that PCA be applied to time series that are stationary. Plot the 
>first column of yieldDat. (You can look at other columns as well. You will see that they are 
>fairly similar.) Does the plot appear stationary? Why or why not? Include your plot with your work.

>Another way to check for stationarity is to run the augmented Dickey-Fuller test. You can do that 
>with the following code:

```{r}
library("tseries")
adf.test(yieldDat[,1])
```

>2.  Based on the augmented Dickey-Fuller test, do you think the first column of yieldDat is 
>stationary? Why or why not?